# Analysis and response

```{r estimates}
d <- readr::read_csv(here::here("data", "data.csv"))
curr_vals <- d %>% 
  filter(year == curr_year)
b <- f(curr_vals$biomass_estimate)
```

## Biomass

Estimates of the California Current Pacific Sardine age-1+ biomass showed a decreasing trend since 2006, reaching historically low levels in recent years [@hill2019]. Declines in recruitment have also occurred since 2005-06, with the exception of a brief period of modest recruitment success from 2009-10 [@hill2019] In particular, the 2011-18 year classes have been among the weakest in recent history. The age-1+ biomass maximum likelihood estimate for the July `r curr_year` forecast is `r b` tonnes [@hill2019].

Few or no sardines have been observed in BC waters during 2013-`r curr_year - 1` in fisheries, surveys or other sources, suggesting curtailed migration and/or stock size. Average estimates of Pacific Sardine trawl catch densities (a catch per unit effort index) from a west coast of Vancouver Island summer pelagic ecosystem night trawl survey in 2006, and 2008-2014 show a decreasing trend from 2006 with no sardines observed the last two years the survey was conducted (2013 and 2014). During the summers of 2015 and 2016 sardines were detected off the west coast of Vancouver Island in trawl catches from other multi-species surveys (i.e. led by DFO or NOAA) but none were detected in 2017 or 2018.

## BC fishery exploitation

```{r calc-tac-landings}
tac <- readr::read_csv(here::here("data", "tac-landings.csv"))
max_bc_catch <- max(tac$bc_catch)
max_bc_catch_year <- tac[which(tac$bc_catch == max_bc_catch),]$year
max_bc_catch <- f(max_bc_catch)

max_tot_catch <- max(tac$total_catch)
max_tot_catch_year <- tac[which(tac$total_catch == max_tot_catch),]$year
max_tot_catch <- f(max_tot_catch)

max_exp <- max(tac$bc_exploitation[!is.na(tac$bc_exploitation)])
max_exp_year <- tac[which(tac$bc_exploitation == max_exp),]$year
max_exp_catch <- f(max_exp)

final_year <- max(tac$year)

# Make catch and tac columns have commas and percent columns have 2 decimal places
tac <- bind_cols(as_tibble(tac[1]),
                 as_tibble(apply(tac[-c(1, 6, 7)], c(1,2), f)),
                 as_tibble(apply(tac[c(6,7)], c(1,2), f, 2)))
tac <- as_tibble(apply(tac, c(1,2), function(x) gsub("NA", "--", x)))
names(tac) <- c("Year",
                "BC TAC",
                "BC Landings",
                "Total Landings",
                "Biomass age-1+",
                "CV age-1+",
                "BC Exploitation (%)")
```

The commercial BC sardine fishery was reinitiated in 2002 following closure since 1947 [@ware1999; @dfo2012]. Most fishing occurred from July to October in association with seasonal sardine migratory behaviour [@dfo2012]. From 2002-2012, the annual total allowable catch (TAC), generally increased as a result of management decisions [@dfo2012]. Prior to 2008, landings were relatively low (less than 5,000 tonnes), but increased considerably from 2008-2012 (up to a maximum of `r max_bc_catch` t in `r max_bc_catch_year`) but were zero in 2013 through to 2018 due to an apparent absence of Pacific Sardine in BC waters and fishery closures in 2015, 2016, 2017, and 2018 (Table \@ref(tab:tac-landings)). Since 2002, total landings of the northern subpopulation
(catches from BC, Washington, Oregon, California and Ensenada Mexico combined) were
highest in `r max_tot_catch_year` at `r max_tot_catch` t. BC annual fishery exploitation rates were estimated as the annual fishery landings divided by the estimated age-1+ biomass in July for each year. Exploitation rate estimates on the stock due to fishing in BC waters show an increase from less than 1% prior to 2009 to a peak in `r max_exp_year` at `r max_exp`%, followed by 0% thereafter (Table \@ref(tab:tac-landings)).

```{r tac-landings}
csasdown::csas_table(tac,
                     align = c('c', 'r', 'r', 'r', 'r', 'r', 'r'),
                     caption = paste0("A summary of recent Pacific Sardine fishery BC TAC, BC landings and total landings of the northern subpopulation for the west coast of North America (BC, Washington, Oregon, California and Ensenada Mexico). Also shown are [@hill2019] estimates of July age-1+ population biomass, coefficients of variation on biomass, and BC exploitation. Total landings for 2002-2005 are from @hill2016 and total landings for 2006-", final_year, " are from @hill2019. The", final_year, " total landings are preliminary. TAC, landings, and biomass values are in metric tonnes."))
```

## Uncertainties

Key uncertainties associated with the 2017 US Pacific Sardine assessment identified in @hill2019 and @punt2018 include information related to:

1. Acoustic species identification, target strength estimation and spatial boundaries associated with the acoustic-trawl survey. Concern over poor near-shore survey coverage has also been expressed by representatives of the U.S and Canadian sardine fishing industry.

2. Empirical weight-at-age data from the fishery and research survey samples and model
characterization of population weight-at-age (e.g. time varying versus time-invariant)

3. Age-length keys to convert acoustic-trawl survey length compositions to age compositions and selectivity parameterization for the acoustic-trawl survey

4. Lack of empirical justification for increasing natural mortality from 0.4 to 0.6 yr-1

5. Uncertainties and concerns identified in past DFO CSAS reviews related to BC Pacific Sardine fishery harvest advice (e.g. @dfo2013) include:

  a. the effect of setting harvest allowances independently of the US and Mexico
  
  b. unknown effects on stock structure and reproductive capacity from fisheries in different regions targeting different age components of the population
  
  c. incidental capture of other species in the sardine fishery
  
  d. the effects of removing sardine from important forage habitat of sardine predators
  
## Harvest options

The July `r curr_year` forecasted age-1+ biomass is `r b` tonnes, which is below the fishery cut-off of 150,000 tonnes. Based on the harvest control rule adopted in 2013, the current recommendation is that there should be no allowable fishery harvest, thus, a TAC of 0 tonnes is recommended for the fishing season starting in `r curr_year`.
